Engineering the Servo Web Browser Engine using Rust

### 2015-04-21 icfp15-servo/response.txt

We would like to thank the reviewers for their time spent reviewing the paper and providing their feedback. We agree with the presentation-related comments, in particular that there is writing that could be improved and we could use a better presentation of Rust, and will incorporate this feedback in a final draft.

In the following, text preceded by a '-' is from a reviewer's question or comment.

Review 1

- Section 4.2, what does it means that Rust embeds serialized generic function
ASTs within "compiled" binary? What's the point? When and how these ASTs
are later used?

TODO: ANSWER

We agree that the problems of JIT correctness and incremental computation are not specific to Rust; they are problems faced by Servo (and all web browser engines, generically).

Review 2

In a final draft, we would be happy to provide a more detailed breakdown of Servo's usage of unsafe Rust code, categorized by purpose.

It is more ambiguous to quantify the costs of Rust's guarnatees. It is much harder to get a program to compile, but debugging basic memory safety errors is largely eliminated. We also did not go into great depth on our usage of Rust's memory management abstractions because we largely use our own. Because Rust does not support integer type specialization (https://github.com/rust-lang/rfcs/pull/884), many of the web-specific optimizations made to optimize default sizes and layouts of data structures based on common web usage patterns require explicit encodings.

### 2015-04-21 icfp15-servo/response.txt

In the following, text preceded by a '-' is from a reviewer's question or comment.

Review 1

- Section 4.2, what does it means that Rust embeds serialized generic function
ASTs within "compiled" binary? What's the point? When and how these ASTs
are later used?

Rust serializes ASTs of items with type parameters into libraries so
that when those items are later instantiated with specific types, the
compiler can generate machine code that is optimized for those
instances. The effect is as with C++ requiring templates to be written
header files. We will clarify.

- Section 5.1, JS JIT, I'm not sure to understand why you consider than JIT
correctness is a problem faced by Rust, inside the Servo implementation.
A little bit of elaboration would have helped me to understand the
problem you want to address here.

### 2015-04-22 icfp15-servo/response.txt

In the following, text preceded by a '-' is from a reviewer's question or comment.

Review 1

- Section 4.2, what does it means that Rust embeds serialized generic function
ASTs within "compiled" binary? What's the point? When and how these ASTs
are later used?

Rust implements generic functions by generating optimized machine code for each
type instantiation. When a library exports a generic function, the compiler
will include a serialized AST in the "compiled" output to support instantiation
by library consumers. We will clarify this in the paper.

Similarly, C++ libraries often contain monomorphic code that is compiled once,
in addition to templates that are exposed in header files. Rust improves
usability by performing this split automatically.

- Section 5.1, JS JIT, I'm not sure to understand why you consider than JIT
correctness is a problem faced by Rust, inside the Servo implementation.
A little bit of elaboration would have helped me to understand the
problem you want to address here.

We agree that JIT correctness is a problem regardless of implementation language. Indeed, we mention it due to the history of security holes resulting from mis-optimizations in SpiderMonkey. The open problem for Servo is how to obtain Rust-like memory-safety guarantees on the JIT compiler's output. We will clarify this in the final draft.

@@

Review 2

In a final draft, we would be happy to provide a more detailed breakdown of Servo's usage of unsafe Rust code, categorized by purpose.

It's harder to quantify the cost/benefit tradeoff of Rust's static checking. Gecko is deployed on hundreds of millions of diverse end-user systems, with the result that Mozilla engineers spend substantial time tracking down individually-rare memory-safety bugs. For this reason, we believe Rust's up-front checking will be especially valuable for a browser engine.

We also did not go into great depth on our usage of Rust's memory management abstractions because we largely use our own. Because Rust does not support integer type specialization (https://github.com/rust-lang/rfcs/pull/884), many of the web-specific optimizations made to optimize default sizes and layouts of data structures based on common web usage patterns require explicit encodings.

### 2015-04-24 icfp15-servo/response.txt

In the following, text preceded by a '-' is from a reviewer's question or comment.

Review 1

- Section 4.2, what does it means that Rust embeds serialized generic function
ASTs within "compiled" binary? What's the point? When and how these ASTs
are later used?

Rust serializes ASTs of items with type parameters into libraries so
that when those items are later instantiated with specific types, the
compiler can generate machine code that is optimized for those
instances (i.e., monomorphized). The effect is as with C++ requiring
templates to be written into header files. We will clarify.

- Section 5.1, JS JIT, I'm not sure to understand why you consider than JIT
correctness is a problem faced by Rust, inside the Servo implementation.
A little bit of elaboration would have helped me to understand the
problem you want to address here.

We agree that the problems of JIT correctness and incremental computation are not specific to Rust; they are problems faced by Servo (and all other web browser engines, generically).

@@

Review 2

In a final draft, we would be happy to provide a more detailed breakdown of Servo's usage of unsafe Rust code, categorized by purpose.

It is more ambiguous to quantify the costs of Rust's guarantees. It is much harder to get a program to compile, but debugging basic memory safety errors is largely eliminated. We also did not go into great depth on our usage of Rust's memory management abstractions because we largely use our own. Because Rust does not support integer type specialization (https://github.com/rust-lang/rfcs/pull/884), many of the web-specific optimizations made to optimize default sizes and layouts of data structures based on common web usage patterns require explicit encodings.

### 2015-04-24 icfp15-servo/response.txt

- Section 4.2, what does it means that Rust embeds serialized generic function
ASTs within "compiled" binary? What's the point? When and how these ASTs
are later used?

Rust implements generic functions by generating optimized machine code for each
type instantiation. When a library exports a generic function, the compiler
will include a serialized AST in the "compiled" output to support instantiation
by library consumers. We will clarify this in the paper.

- Section 5.1, JS JIT, I'm not sure to understand why you consider than JIT
correctness is a problem faced by Rust, inside the Servo implementation.
A little bit of elaboration would have helped me to understand the
problem you want to address here.

We agree that the problems of JIT correctness and incremental computation are not specific to Rust; they are problems faced by Servo (and all other web browser engines, generically).

@@

Review 2

In a final draft, we would be happy to provide a more detailed breakdown of Servo's usage of unsafe Rust code, categorized by purpose.

It's harder to quantify the cost/benefit tradeoff of Rust's static checking. Firefox is deployed on hundreds of millions of diverse end-user systems, with the result that Mozilla engineers spend substantial time tracking down individually-rare memory-safety bugs. For this reason, we believe Rust's up-front checking will be especially valuable for a browser engine.

### 2015-10-11 icse16-servo/intro.tex
Beginning the paper for ICSE16

Servo~\cite{SERVO} is a new web browser engine designed to address the major environment and 
architectural changes over the last decade.
The goal of the Servo project is to produce a browser that enables new applications to be authored
against the web platform that run with more safety, better performance, and better power usage
than in current browsers.
To address memory-related safety issues, we are using a new systems programming language,
Rust~\cite{RUST}.
For parallelism and power, we scale across a wide variety of hardware by building either data-
or task-parallelism, as appropriate, into each part of the web platform.
Additionally, we are improving concurrency by reducing the simultaneous access to data
structures and using a message-passing architecture between components such as the
JavaScript engine and the rendering engine that paints graphics to the screen.
Servo is currently over 400k lines of Rust code and implements enough of the web to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser and its associated libraries.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of programming languages and runtime research.
In this experience report, we discuss the design and architecture of a modern web 
browser engine, show how modern programming language techniques --- many of which
originated in the functional programming community --- address these design 
constraints, and also touch on ongoing challenges and areas of research where we
would welcome additional community input.

>>>>

Servo~\cite{SERVO} is a new web browser engine designed to address the major environment and 
architectural changes over the last decade.
The goal of the Servo project is to produce a browser that enables new applications to be authored
against the web platform that run with more safety, better performance, and better power usage
than in current browsers.
To address memory-related safety issues, we are using a new systems programming language,
Rust~\cite{RUST}.
For parallelism and power, we scale across a wide variety of hardware by building either data-
or task-parallelism, as appropriate, into each part of the web platform.
Additionally, we are improving concurrency by reducing the simultaneous access to data
structures and using a message-passing architecture between components such as the
JavaScript engine and the rendering engine that paints graphics to the screen.
Servo is currently over 400k lines of Rust code and implements enough of the web platform to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser and its associated libraries.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of programming languages and runtime research.
In this experience report, we discuss the design and architecture of a modern web 
browser engine, show how modern programming language techniques --- many of which
originated in the functional programming community --- address these design 
constraints, and also touch on ongoing challenges and areas of research where we
would welcome additional community input.
